EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-20 05_examples.tex
Intermediate commit. Finishing BHZ

The code implementation follows the same guidelines discusses above
for the other examples. There are two specific modifications requied
to the  user. The first is to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone in
order to obtain both the local non-intearcing Hamiltonian
$H_{loc}=\tfrac{1}{N_k}\sum_k H(k)$ and get the local interacting
Green's function
$$
G_{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z) \right]^{-1}
$$
where $\z\in\CCC$, $\Sigma(z)$ is the self-energy function obtained from the DMFT
solution of the problem and $N_k$ is the number of $k$-points. 
The second is to  construct the ``renormalized'' topological
Hamiltonian
$$
H_{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant. $H_{top}$ describes the
low-energy properties of many-body problem and can be used to obtain a 
topological characterization of the interacting system.

\subsubsection{Results}
To capture the possible symmetry breaking in any excitonic channel
driven by the local interaction, we consider the vector order
parameter
$\vec{E}=[E_0,E_x,E_y,E_z]$ where $E_a=\langle\psi^\dagger
\Gamma_{a1}\psi \rangle$.
The first component $E_0$ describes the singlet excitonic state,
whereas the remaing ones correspond to the triplet states with
different spin orientation.   
The analysis of the strong-coupling regime as well as impurity
excitonic susceptibility available in \NAME suggest the possible
instability to an in-plane spin-triplet exciton phase, i.e. with $E_x$
and/or $E_y$ different from zero.
Interestingly, this state breakes several symmetries including
time-reversal and spin SU(2) which protect the topological state.

Thus, here we illustrate the DMFT solution of the excitonic phase as a way to
illustrate the ability of \NAME to capture matter phases with broken
spin-symmetry ({\tt ed\_mode=nonsu2}) and the use of {\tt
  bath\_type=replica}.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !> Set $H_{loc}$ in \NAME
   allocate(Hloc(Nso,Nso))
   Hloc = sum(Hk,dim=3)/Lk
   where(abs(dreal(Hloc))<1d-6)Hloc=zero
   call ed_set_hloc(Hloc)
   !> Setup the replica bath basis using  $M$, $E_0$, $E_z$ and $E_x$ as symmetries
   allocate(lambdasym_vector(Nbath,4))
   allocate(Hsym_basis(Nso,Nso,4))
   Hsym_basis(:,:,1)=Gamma03  ;lambdasym_vector(:,1)= Mh
   Hsym_basis(:,:,2)=Gamma01 ;lambdasym_vector(:,2)= sb_field
   Hsym_basis(:,:,3)=Gamma31 ;lambdasym_vector(:,3)= sb_field
   Hsym_basis(:,:,4)=Gamma11 ;lambdasym_vector(:,4)=-sb_field
   !> Set the replica bath
   call ed_set_Hreplica(Hsym_basis,lambdasym_vector)
   !> Get bath dimension and allocate user bath to this size
   Nb=ed_get_bath_dimension(4)   !There are 4 symmetries
   allocate(Bath(Nb))
   !
   !> Initialize the ED solver (bath is guessed or read from file) 
   call ed_init_solver(bath)
\end{lstlisting}

In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$. 
%
The main effect of the interaction is enclosed in the renormalized
mass term $M_{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real-part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is given proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated with one electron per site, reaching the
conditions for the formation of a Mott insulating state.
In between, at inter-mediate coupling, we observe the formation of a
region of exciton condensation.
In panel (A) of \figu{figEx4} we report the evolution of the order
parameter

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBHZ.pdf}
    \caption{\label{figEx4}%
      \textbf{Topological and Exciton Transition.}
      (a)
    }
\end{figure}

>>>>

The code implementation follows the same guidelines discusses above
for the other examples. There are two specific modifications requied
to the  user. The first is to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone in
order to obtain the local interacting Green's function, used in the
implementation of the self-consistency conditions:
$\GG^{-1}=G_{loc}^{-1}+\Sigma$, with: 
$$
G_{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z)
\right]^{-1}, 
$$
where $\z\in\CCC$  and $\Sigma(z)$ is the self-energy function obtained from the DMFT
solution of the proble.  $N_k$ is the number of $k$-points. 
The second is to  construct the ``renormalized'' topological
Hamiltonian
$$
H_{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant. $H_{top}$ describes the
low-energy properties of many-body problem and can be used to obtain a 
topological characterization of the interacting system.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBHZ.pdf}
    \caption{\label{figEx4}%
      \textbf{Topological and Exciton Transition.}
      (A) Evolution of the spin-triplet,
      in-plane excitonic order parameter $\langle E_x\rangle$ (black solid line) and
      orbital polarization $\langle T_z\rangle$ (orange dashed line) as a function of the
      interaction $U$.
      (B-D) Spectral functions of the two orbital electrons for three
      values of the interaction $U=2.00$ (red circle), $4.00$ (red
      square) and $6.00$ (red diamond) capturing, respectively, the
      QSHI (B), the Excitonic Insulator (C) and the Mott Insulator
      (D). 
    }
\end{figure}

\paragraph{Results}
To capture the possible symmetry breaking in any excitonic channel
driven by the local interaction, we consider the vector order
parameter
$\vec{E}=[E_0,E_x,E_y,E_z]$ where $E_a=\langle\psi^\dagger
\Gamma_{a1}\psi \rangle$.
The first component $E_0$ describes the singlet excitonic state,
whereas the remaing ones correspond to the triplet states with
different spin orientation.   
The analysis of the strong-coupling regime as well as impurity
excitonic susceptibility available in \NAME suggest the possible
instability to an in-plane spin-triplet exciton phase, i.e. with $E_x$
and/or $E_y$ different from zero.
Interestingly, this state breakes several symmetries including
time-reversal and spin SU(2) which protect the topological state.

Thus, here we illustrate the DMFT solution of the excitonic phase as a way to
illustrate the ability of \NAME to capture matter phases with broken
spin-symmetry ({\tt ed\_mode=nonsu2}) and the use of {\tt
  bath\_type=replica}.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !> Set $H_{loc}$
   allocate(Hloc(Nso,Nso))
   Hloc = sum(Hk,dim=3)/Lk
   where(abs(dreal(Hloc))<1d-6)Hloc=zero
   call ed_set_hloc(Hloc)
   !> Setup the replica bath basis:
   allocate(lambdasym_vector(Nbath,4))
   allocate(Hsym_basis(Nso,Nso,4))
   Hsym_basis(:,:,1)=Gamma03  ;lambdasym_vector(:,1)= Mh
   Hsym_basis(:,:,2)=Gamma01 ;lambdasym_vector(:,2)= sb_field
   Hsym_basis(:,:,3)=Gamma31 ;lambdasym_vector(:,3)= sb_field
   Hsym_basis(:,:,4)=Gamma11 ;lambdasym_vector(:,4)=-sb_field
   !> Set the replica bath
   call ed_set_Hreplica(Hsym_basis,lambdasym_vector)
   !> Get bath dimension and allocate user bath to this size
   Nb=ed_get_bath_dimension(4)   !There are 4 symmetries
   allocate(Bath(Nb))
   !
   !> Initialize the ED solver (bath is guessed or read from file) 
   call ed_init_solver(bath)
\end{lstlisting}
where we use the local non-interacting Hamiltonian 
$H_{loc}=\tfrac{1}{N_k}\sum_k H(k)$ to set 
$h^0_{\a\b\s\s'}$, i.e. the impurity Hamiltonian. To anticipate the
possibility of forming an excitonic ordered phase, the replica bath
is constructed out of matrix basis with 4 distinct elements
$\Gamma_{03}$, $\Gamma_{01}$, $\Gamma_{31}$, $\Gamma_{11}$ which
are proportional, respectively, to the mass term, the exciton singlet
$E_0$, the exciton triplet along easy-axis $E_z$ and in-plane
$E_x$ (we assume $E_y=0$ levaraging on residual $U(1)$ in-plane
symmetry.

In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$. 
%
The main effect of the interaction is enclosed in the renormalized
mass term $M_{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real-part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is given proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated with one electron per site, reaching the
conditions for the formation of a Mott insulating state.
As reported in panel (A) of \figu{figEx4}, at intermediate coupling,
the DMFT solutions features the formation of a
region of exciton condensation with $\langle E_x\rangle>0$, $\langle
E_0\rangle=\langle E_z\rangle=0$. 
Unlike the static mean-field description~\cite{Blason2021}, the
transition from the QSHI topological state to the Excitonic Insulator 
(EI) becomes of first-order in presence of quantum fluctuations
corrections~\cite{Continentino,Paoletti} contained in DMFT. On the
other hand, the EI continuously evolves into a Mott Insulator (MI)
for larger interaction (neglecting for simplicity any 
anti-ferromagnetic ordering that would naturally occurr in this
regime).
In the same panel we report the evolution of 
$\langle T_z\rangle$ to illustrate the progressive reduction of the
orbital polarization and, thus, of the effective mass.

To further illustrate the nature of the three distinct phases of this
system, we leverage on the direct access to real-axis spectral
function provided by \NAME solver. In panels (B)-(D) of \figu{figEx4}
we report the spectral functions $-\Im{G}_{a,loc}(\omega)/\pi$ for
$a=1,2$, $\sigma=\up$ and three distinct values of the interaction
$U$ placing the solution, respectively, in the QSHI, the EI and the MI.
The correlated QSHI is characterized by the presence of an 
inverted band gap featuring a substantial orbital spectral mixing. The
EI spectrum is instead characterized by a narrow gap, related to the
finite order parameter, flanked by two sharp resonances and featuring
larger high-energy weight.
Finally, in the MI the two orbital spectral function describes the two 
characteristic Hubbard bands separated by a large Mott gap.

### 2025-05-20 edipack2.tex
Intermediate commit. Adding BHZ nonSU2
Intermediate commit. Finishing BHZ

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack2}\xspace}
\newcommand{\shortrightarrow}{\mathrel{\mkern3mu\rightarrow\mkern}}

@@

\title{EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems}
\author[a,b,c]{L.~Crippa}
\author[a]{I.~Krivenko}
\author[d]{S.~Giuli}
\author[d]{G.~Bellomia}
 \author[c]{A.~Kowalski}
\author[e]{F.~Petocchi}
 \author[f]{M.~Wallerberger}
 \author[g]{G.~Mazza}
 \author[h]{L.~de Medici}
 \author[c,b]{G.~Sangiovanni}
 \author[d,i]{M.~Capone}
\author[i]{A.~Amaricci}
% \author[]{A.~Scazzola}

@@

\address[a]{\ITPHamburg}
\address[b]{\CTQMAT}
\address[c]{\WZBURG}
\address[d]{\SISSA}
\address[e]{\Geneve}
\address[f]{\TUW}
\address[g]{\UPISA}
\address[h]{\ESPCI}
\address[i]{\CNRIOM}
% \address[]{Politecnico di Torino, Turin, Italy}
% \cortext[AA]{Corresponding author:lorenzo.crippa@uni-hamburg.de}
% \cortext[BB] {Corresponding author:giuli@sissa.it}
% \cortext[CC]{Corresponding author:bellomia@sissa.it}

@@

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.
\noindent
{\bf PROGRAM SUMMARY}
\begin{small}
  \noindent
  \\
  {\em Program Title:}  EDIpack2                                        \\
{\em Licensing provisions:} GPLv3\\
{\em Programming language:}  Fortran, Python \\
{\em Classification:} 6.5, 7.4, 20 \\
{\em Required dependencies:} CMake ($>=3.0.0$), Scifortran, MPI\\
{\em Nature of problem:}. \\
{\em Solution method:} .\\
\end{small}

>>>>

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack}\xspace}
\newcommand{\shortrightarrow}{\mathrel{\mkern3mu\rightarrow\mkern}}

@@

\title{EDIpack 2.0: A generic and interoperable Lanczos-based solver for quantum impurity problems}
\author[a,b,c]{L.~Crippa*}
\author[a]{I.~Krivenko*}
\author[d]{S.~Giuli*}
\author[d]{G.~Bellomia*}
\author[c]{A.~Kowalski}
\author[e]{F.~Petocchi}
\author[f]{M.~Wallerberger}
\author[g]{G.~Mazza}
\author[h]{L.~de Medici}
\author[c,b]{G.~Sangiovanni}
\author[d,i]{M.~Capone*}
\author[i]{A.~Amaricci*}
% \author[]{A.~Scazzola}

@@

\address[a]{\ITPHamburg}
\address[b]{\CTQMAT}
\address[c]{\WZBURG}
\address[d]{\SISSA}
\address[e]{\Geneve}
\address[f]{\TUW}
\address[g]{\UPISA}
\address[h]{\ESPCI}
\address[i]{\CNRIOM}
% \address[]{Politecnico di Torino, Turin, Italy}
\cortext[1]{Corresponding authors}
% \cortext[*] {Corresponding author:giuli@sissa.it}
% \cortext[*]{Corresponding author:bellomia@sissa.it}

@@

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.
\noindent
{\bf PROGRAM SUMMARY}
\begin{small}
  \noindent
  \\
  {\em Program Title:}  EDIpack                                        \\
  {\em Licensing provisions:} GPLv3\\
  {\em Programming language:}  Fortran, Python, C/C++\\
  {\em Classification:} 6.5, 7.4, 20 \\
  {\em Required dependencies:} CMake ($>=3.10.0$), Scifortran, MPI\\
  {\em Nature of problem:}. \\
  {\em Solution method:} .\\
\end{small}
